---
title: "Popularity of Justin Bieber and Travis Scott"
author: 
  - Viet Nguyen
  - Doran Wang
  - Yihang Xu
date: October 10, 2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(knitr)
```

# Introduction
This paper analyzes the popularity of songs by Justin Bieber and Travis Scott using data from Spotify. The music industry has shifted dramatically in recent years, with streaming platforms like Spotify playing a key role in determining an artist’s reach and influence. Justin Bieber and Travis Scott represent two distinct genres and musical trajectories, offering a compelling comparison of their popularity based on streaming data.

The goal of this study is to understand how each artist’s songs perform in terms of popularity. We analyze the popularity of individual tracks and look for trends across their discographies. By visualizing and comparing the popularity of songs from these two artists, we aim to identify patterns in their musical success and longevity.

Understanding this data is important for gaining insights into the broader dynamics of the music industry, including factors that contribute to an artist’s sustained popularity. The findings could help predict future trends in music consumption and assist in marketing strategies for artists across genres.

The remainder of this paper is structured as follows: In Section 2, we discuss the data and variables used in the analysis. Section 3 details the modeling approach. Section 4 presents the results, and in Section 5, we conclude with a discussion of the implications of our findings.

# Data
We used the data downloaded from spotify (@spotifyapi) and R (@citeR)
```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
simulated_data <- read_rds(here::here("/Users/nguyenviet/Documents/spotify/data/01-raw_data/justin_bieber_top.rds"))

ggplot(simulated_data, aes(x = reorder(name, -popularity), y = popularity)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  theme_minimal() +
  coord_flip() +  # Flip coordinates for better readability of track names
  labs(title = "Popularity of Tracks", x = "Track Name", y = "Popularity")

```


```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

simulated_data1 <- read_rds(here::here("/Users/nguyenviet/Documents/spotify/data/01-raw_data/travis_scott_top.rds"))

ggplot(simulated_data1, aes(x = reorder(name, -popularity), y = popularity)) +
  geom_bar(stat = "identity", fill = "pink", color = "black") +
  theme_minimal() +
  coord_flip() +  # Flip coordinates for better readability of track names
  labs(title = "Popularity of Tracks", x = "Track Name", y = "Popularity")
```


```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
simulated_data2 <- read_rds(here::here("/Users/nguyenviet/Documents/spotify/data/01-raw_data/travis_scott.rds"))
simulated_data2 <- simulated_data2 %>% mutate(artist_name = "Travis Scott")

simulated_data3 <- read_rds(here::here("/Users/nguyenviet/Documents/spotify/data/01-raw_data/justin_bieber.rds"))
simulated_data3 <- simulated_data3 %>% mutate(artist_name = "Justin Bieber")

rbind(simulated_data2, simulated_data3) |>
  select(artist_name, release_date, total_tracks) |>
  mutate(release_date = ymd(release_date)) |> 
  ggplot(aes( x = release_date, y = total_tracks, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "release date",
    y = "total_tracks",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

# Discussions
The analysis of Justin Bieber’s and Travis Scott’s popularity on Spotify reveals several key insights. First, Bieber’s music, spanning multiple albums and genres, shows consistent high popularity across a wide range of tracks. His collaboration-heavy tracks, particularly those featuring well-known artists, tend to rank highly, suggesting that collaborations may play a significant role in his sustained success. Meanwhile, Travis Scott’s discography reflects a distinct pattern, with a sharp increase in track popularity following the release of his more recent albums, indicating his growing prominence in the hip-hop genre.

The differences between Bieber and Scott’s popularity trends can be attributed to several factors, including their genre, target audience, and marketing strategies. Bieber’s mainstream appeal across pop and R&B genres contrasts with Scott’s niche within the hip-hop community, but Scott’s rising influence demonstrates how genre specialization can also lead to mainstream success. Additionally, release timing and the number of tracks per album appear to impact their overall popularity trends.


# Appendix
These are the sample variables we used from the data we downloaded.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
out <- simulated_data[c("name","popularity")]
out2 <- simulated_data1[c("name","popularity")]
out3 <- simulated_data2[c("total_tracks","release_date")]
out4 <- simulated_data3[c("total_tracks","release_date")]

kable(head(out))
kable(head(out2))
kable(head(out3))
kable(head(out4))

```

# References

